---
title: Programming in dplyr - Row Filtering
author: Thomas Jc
date: '2018-10-28'
slug: programming-in-dplyr-row-filtering
categories:
  - rstats
tags:
  - dplyr
  - tidyeval
editor_options: 
  chunk_output_type: console
---

#### How do you program the row filtering dynamically when working with `dplyr` verb?

Say we have a dataframe with 2 columns, x and y.

```{r, message=FALSE}
library(tidyverse)
library(rlang)

# make some sample
set.seed(1234)
df <- 2 %>% 
    rerun(rnorm(100, 10, 2) %>% round()) %>% 
    bind_cols()
colnames(df) <- c("x", "y")

# simple plot
p <- df %>% 
    ggplot(aes(x, y)) + 
    geom_point(position = position_jitter(width = 1), 
               alpha = 0.7) +
    theme_minimal()
p
```

**Our objective here is to 'zone' out the points that fall in the 50th percentile of both x and y.**

```{r}
summary(df)

# filter values within (w) q1, q3
(w <- df %>% 
    filter(x >= quantile(x, probs = 0.25),
           x <= quantile(x, probs = 0.75),
           y >= quantile(y, probs = 0.25),
           y <= quantile(y, probs = 0.75)))

# points that fall outside q1, q3
v <- anti_join(df, w, by = c("x", "y"))

# modify previous plot
w %>% 
    ggplot(aes(x, y)) + 
    geom_jitter(width = 1, col = "red") + 
    geom_jitter(data = v, aes(x, y), alpha = 0.7) + 
    theme_minimal()
```

The way that we express our filtering conditions is what we aim to simplify in this post.

```{r, eval=FALSE}
filter(
    data = df,
    x >= quantile(x, probs = 0.25),
    x <= quantile(x, probs = 0.75),
    y >= quantile(y, probs = 0.25),
    y <= quantile(y, probs = 0.75)
    ))
```

## Expression

We start from forming a simple expression in literal text. 

```{r}
text_expression <- function(var, ops, pct) {
    sprintf("%s %s quantile(%s, probs = %g)", var, ops, var, pct)
}
text_expression("x", ">=", .75)
text_expression("y", "<=", .75)
```

## Rlang

We use `rlang` to transform text into R expression.

```{r}
custom_filter <- function(x, filter_expr) {
    expr <- parse_expr(filter_expr)
    df %>% filter(!!expr)
}
custom_filter(x, text_expression("x", ">=", .75))
```

## Multiple Arguments

It takes the following steps, 

1. Use `purrr::map` to form multiple expressions
2. Use `;` or `\n` to join formed expressions
3. Use `rlang::parse_exprs` to transform texts into R expressions
4. Use `!!!` (3x bang) to slice expressions

```{r}
# the following has 3 expressions
parse_exprs("NULL; list()\n foo(bar)")
```

Let's continue onto our example. 

```{r}
# form 'head' of our expressions
(ehead <- map(.x = list("x", "y"), .f = paste, list(">=", "<=")))

# form 'taik' of our expressions
(etail <- map(.x = list("x", "y"), 
             .f = ~ sprintf('quantile(%s, probs = %g)', .x, .y),
             list = c(.25, .75)))

# combine head and tail
(exprs <- map2(ehead, etail, ~ sprintf("%s %s", .x, .y)))

# collapse our expressions into a long string
(exprs <- exprs %>% 
    unlist() %>% 
    # use ';' or "\n"
    paste0(collapse = ";"))
```

## Apply Custom Filtering

```{r}
# modify our custom filtering to fit multiple expressions
custom_filter = function(x, exprs) {
    # take note of the plurar form
    exprs_list <- parse_exprs(exprs)
    # apply on data frame
    x %>% filter(!!!exprs_list)
}
# thats it
(w2 <- df %>% custom_filter(exprs))
```

```{r, message=FALSE}
w2 %>% 
    ggplot(aes(x, y)) +
    geom_point(position = position_jitter(width = 1), col = "red") + 
    geom_point(data = anti_join(df, w), aes(x, y)) +
    theme_minimal()

# final check
all(w == w2)
```

