---
title: Data.Table
description: Performing long-to-wide transformation on large data frame
author: Thomas Jc
date: '2018-02-09'
slug: long-to-wide
categories:
  - rstats
tags:
  - data.table
---


A quick walkthrough on using ``data.table`` in performing long-to-wide operation on data frame. Significant improvement on speed.

```{r, message=FALSE}
library(dplyr)
library(tidyr)
library(data.table)

# set
options(DT.options = list(pageLength = 5, dom = 'tip'))
set.seed(123)
rand <- runif(1e7, min = 1, max = length(LETTERS)) %>% round()

# Make some dummy data
df <- data.frame(
    letters = LETTERS[rand] %>% sample(),
    num = (1:length(LETTERS))[rand] %>% sample()
)
# Add count for spreading later
df <- df %>% group_by(letters) %>% mutate(count = 1:length(letters)) %>% arrange(letters, count)
head(df, 10)
```

Let's try out different methods on converting above long data frame to wide format. Our objective is to group them by `letters` and each row lists how many times the particular letter is called, from 1 to N. 

```{r}
# For example
(df %>% spread(count, num) %>% head())[1:5, 1:10]
```

### Method 1: The Usual Spread

```{r, cache=TRUE}
t1 <- system.time(m1 <- df %>% spread(count, num))
(m1 %>% head())[1:5, 1:10]
```


### Method 2: D-Cast


```{r, cache=TRUE}
t2 <- system.time(m2 <- df %>% dcast(letters ~ count, value.var = "num"))
(m2 %>% head())[1:5, 1:10]
```


### Method 3: D-Cast with Data Table

```{r}
# Convert to data table
dt <- df %>% as.data.table()
t3 <- system.time(m3 <- dt %>% dcast(letters ~ count, value.var = "num"))
(m3 %>% head())[1:5, 1:10]
```

### Result 

Shown below.

```{r}
res <- list(t1, t2, t3)
names(res) <- c("Spread", "DCast", "DCast w/ Data Table")
res
```

