---
title: Programming in dplyr - Column Manipulation
author: Thomas Jc
date: '2018-09-07'
slug: programming-in-dplyr
categories:
  - rstats
tags:
  - dplyr
  - tidyeval
editor_options: 
  chunk_output_type: console
---

#### How do you program the column manipulation dynamically when working with `dplyr` verb? 

Say, we have a dataframe with multiple columns in pair,

```{r, message=FALSE}
library(tidyverse)

# simulate some data
df <- rerun(3 * 2, rnorm(10, 5, 2)) %>% bind_cols()
# imagine some column names
colnames(df) <- map(c("weight", "height", "length"), paste0, c(".x", ".y")) %>% unlist()
df
```

and we try to calculate the difference between metrices *.x* and *.y*

## Expression

Let formulate a form of "expression", 

```{r}
calc_diff <- function(var) {
    # the desire expression is (var.x - var.y)
    sprintf("%s.x - %s.y", var, var)
}
# example
calc_diff("weight")
```

## Rlang

So we get our desire expression, but how do we pass them to `dplyr` to understand? We can use `rlang` to transform text into R expressions.

```{r}
calc_diff <- function(df, var) {
    # parse as R code
    expr <- sprintf("%s.x - %s.y", var, var) %>% 
        rlang::parse_expr()
    df %>% mutate(!!expr)
}

df %>% 
    # add new column
    calc_diff("weight") %>% 
    # focus on result 
    select(contains("weight"))
```

## Column Name

Well, we need a way to add name to the new column. 

```{r}
calc_diff <- function(df, var) {
    # to transform a quosure to a string 
    new_col_name <- quo_name(paste0(var, ".diff"))
    expr <- sprintf("%s.x - %s.y", var, var) %>% 
        rlang::parse_expr()
    # so that we can use it in dplyr context
    df %>% mutate(!!new_col_name := !!expr)
}

df %>% 
    # it works!
    calc_diff("weight") %>% 
    select(contains("weight"))
```

## Many Columns

So now we know how to do with 1 variable, how do we apply it to many? 

1. Use `map` as per normal to loop and generate multiple expressions
2. Use `set_names` to give expressions list a name
3. Use `!!!` (3x *!* not 2) to splice multiple arguments in `mutate`. It will assign new variable names using names given by list (in step 2).

```{r}
calc_diff <- function(df, ...) {
    # first, an expressions list
    n_exprs <- 
        # loop over multiple inputs
        map(list(...), ~ sprintf("%s.x - %s.y", .x, .x) %>% 
                       rlang::parse_expr()) %>% 
        set_names(map(list(...), ~ paste0(.x, ".diff")))
    
    # splices above expressions
    df %>% mutate(!!!n_exprs)
}

# tada~~
df %>% 
    calc_diff("weight", "height", "length")
```





