---
title: parDist
description: Compute Distance Matrix for Large Object
author: Thomas Jc
date: '2018-02-09'
slug: par-dist
categories:
  - rstats
tags:
  - parallelDist
---

### Problem Statement:

> We have a large dataset. We need to compute its distance matrix (i.e. for clustering purposes). The complexity for a N \* P matrix is `N(N-1)/2 * 3P`


### Create Sample

```{r, cache=TRUE}
library(dplyr)

pts <- rnorm(1e4, mean = c(5, 5, 10, 10, 15, 15), sd = 1) %>% matrix(ncol = 2, byrow = T)
pts %>% plot(xlab = "x", ylab = "y")
```

### Package ‘parallelDist’

``parallelDist`` is a fast parallelized alternative to R's native ``dist`` function. The package is mainly implemented in C++ and leverages the ``RcppParellel`` package to parallelize the distance computations. In addition, it also uses `Armadillo` linear algebra library to optimize matrix operations during distance calculations. 

In short, to compute distance matrix for large data object, use ``parallelDist`` because it is much faster.


### Demo

Say we wish to compute distance matrix to compute ``silhouette`` distance for sample points above to determine the optimal number of cluster (which we already knew is 3).

```{r, cache=TRUE}
# By default, parDist returns a dist object
# Here we convert to matrix for minor efficiency
dist.euclidean <- parallelDist::parDist(pts, method = "euclidean") %>% as.matrix()

# A custom function for looping
compare_silhouette <- function(k){
    kmeans(pts, centers = k, nstart = 20, iter.max = 50)$cluster %>% 
        # use dmatrix instead of dist
        cluster::silhouette(dmatrix = dist.euclidean) %>% 
        summary() %>% 
        # extract avergae silhouette width
        `[[`('avg.width')
}

# Here we try out various number of clusters
res <- lapply(2:5, compare_silhouette)

library(ggplot2)
data.frame(x = 2:5, y = unlist(res)) %>% ggplot(aes(x, y)) + geom_line() + geom_point() + theme_light()
```




