---
title: Bar, Bended Bar, and Treemap
description: Visualizing Count Observations
author: Thomas Jc
date: '2018-01-21'
slug: visualize-count-observations
categories:
  - rstats
tags:
  - ggplot2
  - drivethru
---

In this exercise, we will visualize Standard & Poor's 500 Index industry composition.

### Gettting sp500_sectora

We will scrap the sp500_sectora from [Wikipedia page](https://en.wikipedia.org/wiki/S%26P_500_Index), with `rvest`.

```{r, message=FALSE}
library(dplyr)
library(rvest)

url <- "https://en.wikipedia.org/wiki/List_of_S%26P_500_companies"

sp500 <- url %>% 
    read_html() %>% 
    # How do I select css? 
    html_node(css = "table.wikitable") %>% 
    html_table() %>% 
    mutate(sector = factor(`GICS Sector`)) %>% 
    as_tibble()
sp500
```

<small>Note: Write a post about `rvest` node selector.</small>

### Pie Chart

Let's begin with an ~~awful~~ pie chart.

```{r}
library(ggplot2)
library(RColorBrewer)

# Set ggplot theme
old <- theme_set(theme_light() + theme(legend.position = "bottom"))

# Set colour palette for repeated use
pal <- scale_fill_brewer(
    palette = "Set3",
    guide = guide_legend(
    title = "Sector",
    title.position = "top",
    label.position = "top",
    keyheight = 0.5,
    ncol = 4
    )
    )

# Pie Chart
sp500 %>% 
    ggplot(aes(factor(1), fill = sector)) + 
    geom_bar(width = 1) + 
    # Pie chart is simply a change in the polar coordinate
    coord_polar(theta = "y") +
    labs(x = "", y = "Count") +
    pal
```


The disadvantage of using a pie chart is that we cannot clearly visualize the differences, e.g. how much difference is there between the biggest and second biggest industry?

### Bar Chart

We should be quite familiar with this. 

```{r}
# Summarize sector by count
sp500_sector <- sp500 %>% group_by(sector) %>% summarise(n = n())

# Normal Bar Chart 
sp500_sector %>% 
    ggplot(aes(reorder(sector, -n), n, fill = sector)) + 
    geom_bar(stat = "identity") +
    # We hide axis x text 
    theme(axis.text.x = element_blank()) +
    # and manually replace with geom_text
    geom_text(aes(y =  n/2, label = sector), 
              show.legend = FALSE, 
              angle = 90, 
              size = 3, 
              nudge_y = +20) +
    labs(x = "", y = "Count") +
    pal
```


### Polar Bar Chart

What if we bend the bars? (Inspired by [Apple Watchface](https://9to5mac.files.wordpress.com/2016/06/img_5044.png?w=312))

```{r}
# The Circular Bar Chart Way
sp500_sector %>% 
    ggplot(aes(reorder(sector, n), n, fill = sector)) + 
    geom_bar(stat = "identity") + 
    # Expand the breaks & limit to look nicer
    scale_y_continuous(breaks = scales::pretty_breaks(10), 
                       expand = c(0, 0.8)) +
    coord_polar(theta = "y", direction = 1) +
    labs(x = "", y = "Count") +
    pal
```

The main benefit is that we get a more organized display of sorted ranking. Not bad.

### Treemap

In my opinion, Treemap is good when you wish to zoom into a particular group of objects. In our case, attention is drawn to the biggest industries. Treemap is also space-efficient. Legend and axis labels are often expendable. 

```{r}
library(treemapify)

sp500_sector %>% 
    mutate(pct = scales::percent(n / sum(n))) %>% 
    ggplot(aes(area = n, fill = sector, label = paste(sector, "\n", pct))) + 
    geom_treemap(show.legend = FALSE) + 
    geom_treemap_text(place = "centre", size = 9) + 
    pal
```

That's it. A quick 10. Till next time.
