---
title: Is James Harden been given special treatment?
author: tmasjc
date: '2019-10-06'
slug: is-james-harden-been-given-special-treatment-by-referee
categories: []
tags: 
    - isittrue
---

This video ["8 Unbelievable Charts That Will Change The Way You See The NBA"](https://www.youtube.com/watch?v=VCIKek1f-Hs) pops out on my Youtube homepage. The first chart he argues that James Harden has been given way more free throws (834) by NBA referees compare to the rest of the players. I find the chart *'unbelievable'* indeed and so I decided to look up some data on [basketball-reference](https://www.basketball-reference.com) to determine its validatity. 

Conclusion, not true. 

```{r, warning=FALSE, message=FALSE, cache=TRUE}
library(tidyverse)
library(ggthemes)
library(rvest)
library(janitor)
library(plotly)
# ggplot theme
old <- theme_set(theme_minimal() + theme(text = element_text(family = 'Menlo')))

# we scrap the data for the past 10 years
yrs  <- c(2008:2018)
urls <- str_glue("https://www.basketball-reference.com/leagues/NBA_{yrs}_per_poss.html")
tbls <- map(urls, {
    . %>% 
        read_html() %>% 
        html_node("#per_poss_stats") %>% 
        html_table()
}) %>% set_names(yrs)

# clean up a little
dat <- map(tbls, ~ which(nchar(names(.)) >= 1)) %>% 
    map2(.x = tbls, .y = ., .f = ~ .x[.y]) %>% 
    bind_rows(.id = "YEAR") %>% 
    as_tibble() %>% 
    clean_names() 

df <- dat %>% 
    filter(player != "Player", tm != "TOT") %>% 
    # these are the useful columns to us 
    # for data glossary, refer to basketball-reference.com
    select(year, player, tm, pos, g, fta) %>% 
    mutate(pos = str_extract(pos, "\\w*"),
           g   = as.integer(g),
           fta = as.double(fta))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggthemes)
library(rvest)
library(janitor)
library(plotly)
# ggplot theme
old <- theme_set(theme_minimal() + theme(text = element_text(family = 'Menlo')))
```


```{r, fig.width=7}
set.seed(1212)
p <- df %>% 
    # filter those who completed at least 60 percent or 50 games of a season
    filter(g >= 50) %>% 
    group_by(year) %>% 
    top_n(10, fta) %>% 
    arrange(desc(fta), .by_group = TRUE) %>% 
    # filter(year == "2017")
    ggplot(aes(year, fta, 
               text = str_glue("Name: {player}
                               Team: {tm}
                               Games: {g}
                               Position: {pos}
                               fta: {fta}"))) + 
    geom_point(aes(col = pos), alpha = .7, size = 3,
               position = position_jitter(width = .1, height = .3)) + 
    labs(x     = "",
         y     = "",
         title = "Free Throw Attempts per 100 Team Possessions",
         col   = "Position") 

ggplotly(p, tooltip = "text") %>% 
    layout(legend = list(orientation = "h", x = 0, y = -.1))
```




