---
title: 'Simulate Tossing Coin Game'
author: tmasjc
date: '2019-02-24'
slug: tossing-coin
categories: []
tags: 
    - monte-carlo
editor_options: 
  chunk_output_type: console
---

The use of simulation experiments to better probability patterns is called the Monte Carlo method. This is an example.

#### Tossing Coin Game

Tom and Jerry play a simple game of tossing a coin. If given heads, Tom pays Jerry a dollar. Otherwise, Jerry pays Tom a dollar. 

```{r}
set.seed(1212)

# how many tosses?
number_of_tosses = 50

# simulate n tosses
# 1 = head, -1 = tail
(tosses <- sample(c(1, -1), size = number_of_tosses, replace = TRUE, prob = c(.5, .5)))
```

Let's observe some games' progression, from the perspective of Tom.

```{r, message=FALSE}
library(purrr)

# convert game to a function
new_game <- function(n = 1) {
  tosses = sample(c(1, -1),
                   size = n,
                   replace = TRUE,
                   prob = c(.5, .5))
}

# progress of 4 different games
par(mfrow = c(2, 2))
walk(1:4, ~ {
  plot(cumsum(new_game(number_of_tosses)), type = "l", ylab = "Winnings")  
  abline(h = 0, col = "red", lty = 3)
})
```

Let's calculate the winnings distribution.

```{r, message=FALSE}
library(dplyr)
library(ggplot2)
library(ggthemes)

old <- theme_set(theme_tufte() + theme(text = element_text(family = "Menlo")))

# modify function above
new_game <- function(n = 1) {
  tosses = sample(c(1, -1),
                   size = n,
                   replace = TRUE,
                   prob = c(.5, .5))
  # return winnings
  sum(tosses)
}
# repeat many times
reps <- 10e3 %>% rerun(new_game(number_of_tosses)) %>% unlist()

# bind into data frame
reps %>%
    as_data_frame() %>%
    ggplot(aes(value)) +
    geom_histogram(binwidth = 1) +
    scale_x_discrete(limits = seq(-24, 24, 4))
```

*Question: Why there isn't any odd number?*

We can also calculate the chances where Tom break even, either from simulation result above, or using binomial distribution. They are pretty close.

```{r}
# how many breakeven
length(which(reps == 0)) / length(reps)

# approx. form using binomial distribution
dbinom(25, size = 50, prob = 0.5) 
```

What is the likely number of tosses Peter will be winning in a game of 50 tosses? What is the likely maximum fortune Peter will be winning in a game of 50 tosses?

```{r, message=FALSE}
library(tidyr)

# modify function above
new_game <- function(n = 1) {
  tosses = sample(c(1, -1),
                   size = n,
                   replace = TRUE,
                   prob = c(.5, .5))
  # track progress
  cum.win = cumsum(tosses)
  # return
  list(final = sum(tosses), 
    leads = sum(cum.win > 0), 
    max = max(cum.win))
}
new_game(number_of_tosses)

# repeat
set.seed(1234)
reps <- 10e3 %>% rerun(new_game(number_of_tosses))

reps %>%
    bind_rows() %>%
    # collapse columns into one
    gather(var, val) %>%
    ggplot(aes(val)) +
    geom_histogram(binwidth = 1) +
    facet_wrap(~ var, ncol = 1,
               labeller = labeller(
                   var = c(final = "Final Winnings",
                           leads = "# Leadings",
                           max = "Max Fortunate")
               )) +
    labs(x = "", y  = "")
```



