---
title: 'Exercise: Sales Forecast using MCM'
author: tmasjc
date: '2019-04-09'
slug: sales-forecast-using-monte-carlo
categories: []
tags: 
    - monte-carlo
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## Background

Often, Monte Carlo simulation can come in handy to calculate risk or evaluate investments in projects. This is a simple demonstration.

## Exercise

The following provides the breakdown of profit made by a business unit. All metrics are measured in daily basis.

    Profit = Income - Expenses
    
        Income = Sales (S) * Profit Margin per Sale (M)
        
            M assumes an uniform dist. from $350 to $400
        
            S = Number of Leads (L) * Conversion Rate (R)
        
                 L assumes an uniform dist. with from 3000 to 4000
                 
                 R assumes a normal dist. with mean of 4% and sd of 0.5%
            
        Expenses = Fixed Overhead (H) + Total Cost of the Leads (C)
        
            C = Cost Per Lead (Cpl) * Number of Leads (L)
            
                Cpl assumes an uniform dist. from $8 to $10
        
            H assumes a constant of $20000
            

In summary, 


    Profit = Leads * Conversion Rate * Profit Margin per Sale - (Cost per Lead * Leads + Fixed Overhead)

            
```{r lib, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggthemes)

# ggplot theme
old <- theme_set(theme_tufte() + theme(text = element_text(family = 'Menlo')))
options(scipen = 99)
```

## Profit Forecast Model

An oversimplified daily profit forecast model,

```{r forecast_profit}
# profit forecast function
forecast_profit <- function() {
    
    # number of leads
    L <- runif(n = 1, min = 3000, max = 4000)
    
    # conversion rate
    R <- rnorm(n = 1, mean = .04, sd = .005)
    
    # average profit margin per sale
    M <- runif(n = 1, min = 350, max = 400)
    
    # average cost per lead
    Cpl <- runif(n = 1, min = 8, max = 10)
    
    # fixed overhead
    H <- 20e3
    
    # return profit
    (L * R * M) - (L * Cpl + H)
}

# run simulation 
set.seed(1212)
nn = 10e3
daily <- tibble(ind = 1:nn, profit = replicate(nn, forecast_profit()))
```

```{r plot_outcome_hist}
# view outcome distribution
daily %>% 
    ggplot(aes(profit)) +
    geom_histogram(aes(y = ..density..), 
                   col = "black", fill = "salmon", alpha = .7, bins = 45) + 
    geom_density(col = "red", size = 1.5) +
    geom_vline(xintercept = quantile(daily$profit, c(0.025, 0.5, 0.975)), 
               lty = 4, col = "navyblue", size = 1.2) + 
    scale_x_continuous(labels = scales::dollar) +
    theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
    labs(x = "Profit", y = "Density")
```

If we set a profitability goal of **`r scales::dollar(100e3)` a month**, what is the probability that we achieve that? How about the probability that we lose money?

```{r forecast_monthly}
# forecast monthly goal and risk
forecast_monthly <- function(df, n = 1000, goal = 100e3) {
  
  # we draw samples from our daily forecast 
  res <- replicate(n, sum(sample(df$profit, 30)))
  
  # likelihood that we exceed our profit goal
  p1 <- length(which(res >= goal)) / n
  
  # likelihood that we do not make profit (incur loss)
  p2 <- length(which(res < 0)) / n
  
  return(
    list(
      df = tibble(ind = 1:n, profit = res),
      goal = goal,
      profit_prob = p1,
      loss_prob = p2
    ))
  
}
```

```{r}
daily <- tibble(ind = 1:nn, profit = replicate(nn, forecast_profit()))
```

```{r print_monthly}
monthly <- forecast_monthly(daily)
sprintf("Probability of hitting goal is %s", scales::percent(monthly$profit_prob))
sprintf("Probability of incurring losses is %s", scales::percent(monthly$loss_prob))
```

We can also plot the cumulative probability for clearer visualization.

```{r plot_cumulative}
# plot cumulative probability
plot_cumulative <- function(daily, ...) {
  monthly <- forecast_monthly(daily, ...)
  monthly$df %>% 
    ggplot(aes(profit)) + 
    stat_ecdf() + 
    geom_vline(xintercept = c(0, monthly$goal), lty = 4, col = "lightgray") +
    geom_hline(yintercept = c(1 - monthly$profit_prob, monthly$loss_prob), lty = 4, col = "lightgray") +
    annotate(x = c(monthly$goal, 0), 
             y = c(1 - monthly$profit_prob, monthly$loss_prob), 
             geom = "point", col = "red", size = 2) +
    scale_x_continuous(labels = scales::dollar) +
    scale_y_continuous(labels = scales::percent) + 
    labs(x = "Profit", y = "Cumulative Probability")
}
plot_cumulative(daily)
```

## Update Model

What if we further assume that cost per lead and conversion rate are correlated?

```{r cost_conversion_corr}
# assume the true parameter (conversion) is some cubic function
x <- seq(8, 10, by = .1)
y = 9.46 - 3.206667 * x + 0.36 * x ^ 2 - 0.01333333 * x ^ 3

# bind to data frame
cc <- tibble(x, y) %>% 
    distinct() %>% 
    arrange(x) %>% 
    rename(cost_per_lead = x, conversion = y)

# if we take linear regression for curve fitting
fit <- lm(y ~ x, data = cc)
```


```{r update_forecast}
# update profit function
update_forecast <- function() {
    
    # number of leads
    L <- runif(n = 1, min = 3000, max = 4000)
    
    # average profit margin per sale
    M <- runif(n = 1, min = 350, max = 400)
    
    # average cost per lead
    Cpl <- runif(n = 1, min = 8, max = 10)
    
    # conversion rate
    #R <- 9.46 - 3.206667 * Cpl + 0.36 * Cpl ^ 2 - 0.01333333 * Cpl ^ 3
    R <- predict(fit, newdata = data.frame(x = Cpl))
    
    # fixed overhead
    H <- 20e3
    
    # return profit
    (L * R * M) - (L * Cpl + H)
}
# remember to update function name
daily <- tibble(ind = 1:nn, profit = replicate(nn, update_forecast()))
```

```{r ref.label="print_monthly"}

```

## Sensitivity Analysis

What if we are offered an option to increase our leads at the cost of fixed overheads increase? 

```{r sens_forecast}
# add variation to leads and overhead
sens_forecast <- function() {
    
    # number of leads
    L <- runif(n = 1, min = 3000, max = 4000) * runif(1, min = 1.15, max = 1.25)
    
    # average profit margin per sale
    M <- runif(n = 1, min = 350, max = 400)
    
    # average cost per lead
    Cpl <- runif(n = 1, min = 8, max = 10)
    
    # conversion rate
    #R <- 9.46 - 3.206667 * Cpl + 0.36 * Cpl ^ 2 - 0.01333333 * Cpl ^ 3
    R <- predict(fit, newdata = data.frame(x = Cpl))
    
    # fixed overhead
    H <- 20e3 * runif(1, min = 1.25, max = 1.30)
    
    # return profit
    (L * R * M) - (L * Cpl + H)
}
# 
daily <- tibble(ind = 1:nn, profit = replicate(nn, sens_forecast()))
```

```{r ref.label="print_monthly"}

```


## Finding Optimal

What is the maximum cost per lead we can accept if we wish to cover our probability of losses at X%? 

```{r, cache=TRUE}
# profit forecast function
reduce_risk <- function(max) {
    
    # number of leads
    L <- runif(n = 1, min = 3000, max = 4000)
    
    # conversion rate
    R <- rnorm(n = 1, mean = .04, sd = .005)
    
    # average profit margin per sale
    M <- runif(n = 1, min = 350, max = 400)
    
    # fixed overhead
    H <- 20e3
    
    # we need to vary this
    Cpl <- runif(n = 1, min = 8, max = max)
    
    # return profit
    (L * R * M) - (L * Cpl + H)
}

# grid linear search
cpl <- seq(8.1, 9.9, by = 0.05)

#-- we increase the number of trials here **
opt <- purrr::map_lgl(cpl, ~ {
  d = replicate(nn * 2, reduce_risk(.x))
  m = forecast_monthly(data.frame(profit = d), nn * 2)
  m$loss_prob > 0.05
}) %>% set_names(cpl)

# the first occurance of 'true'
sprintf("Maximum cost per lead allowed to reduce risk down to 0.05 is $%s", names(which.max(opt) - 1))
```




