---
title: Biased Coin Toss
author: Thomas Jc
date: '2019-04-21'
slug: biased-coin-toss
categories: []
tags: 
    - monte-carlo
---

From this paper "[What's Past Is Not Prologue](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3034686)" comes a quiz, 

> You are presented with two coins: one is fair, and the other has a 60% chance of coming up heads. Unfortunately, you don’t know which is which. How many flips would you need to perform in parallel on the two
coins to give yourself a 95% chance of correctly identifying the biased one?

The following demonstrates 2 simulation methods to obtain the result. 

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggthemes)
library(discreteRV)
set.seed(1212)

# ggplot theme
old <- theme_set(theme_tufte() + theme(text = element_text(family = 'Menlo')))
```

```{r, cache=TRUE}
# limit range from 50 to 200 flips
trials <- 50:200

# Method One --------------------------------------------------------------


# actually simulate the action of tossing 2 coins
method_one <- function(x) {
    regular = rbinom(x, 1, .5)
    biased = rbinom(x, 1, .6)
    # return
    sum(biased) - sum(regular)
}

# repeat action and bind into a data frame
res.one <- map_dbl(trials, ~ {
    sum(replicate(3000, method_one(.x)) > 0) / 3000
}) %>% bind_cols(ind = trials, prob = .)

# first occurance
(min.flip.one <- filter(res.one, prob >= 0.95)[1, ])

# vis result
res.one %>% 
    ggplot(aes(ind, prob)) +
    geom_line() + 
    geom_vline(xintercept = min.flip.one$ind, lty = 4, col = "red") +
    geom_hline(yintercept = min.flip.one$prob, lty = 4, col = "red") +
    annotate("text", x = min.flip.one$ind, y = .5, label = as.character(min.flip.one$ind)) + 
    scale_y_continuous(breaks = seq(.5, 1, .1), labels = scales::percent)


# Method Two --------------------------------------------------------------


# we will reuse this function repeatedly to test different diff trials
method_two <- function(x) {
    # how many tosses?
    regular <- RV(1:x, dbinom(1:x, size = x, prob = .5))
    biased <- RV(1:x, dbinom(1:x, size = x, prob = .6))
    
    # joint pmf
    j2 <- joint(regular, biased)
    
    # probability here can be understood as proportion of combination of possible outcomes
    tibble(prob = probs(j2), comb = outcomes(j2)) %>%
        separate(comb, c("fair", "biased"), ",", convert = TRUE) %>% 
        # if biased lands more than unbiased coin, choose correctly
        # if both lands the same, choose randomly
        mutate(correct = case_when(
            biased > fair ~ 1,
            biased == fair ~ 0.5,
            biased < fair ~ 0
        )) %>% 
        # the likelihood that guessing correctly happens
        mutate(likelihood = prob * correct) %>% 
        filter(correct > 0) %>%
        summarise(conf = sum(likelihood)) %>% 
        pull(conf)
}

# loop thru trials
res.two <- map_dbl(trials, ~ method_two(.x)) %>% bind_cols(ind = trials, prob = .)

# minimum flip to obtain 95% confidence
(min.flip.two <- filter(res.two, prob >= 0.95)[1, ])

# vis result
res.two %>% 
    ggplot(aes(ind, prob)) + 
    geom_line() + 
    geom_vline(xintercept = min.flip.two$ind, lty = 4, col = "red") +
    geom_hline(yintercept = min.flip.two$prob, lty = 4, col = "red") +
    annotate("text", x = min.flip.two$ind, y = .5, label = as.character(min.flip.two$ind)) + 
    scale_y_continuous(breaks = seq(.5, 1, .1), labels = scales::percent)
```




