---
title: In A Village Where People Only Want Boys
author: Thomas Jc
date: '2020-02-29'
slug: in-a-village-where-people-only-want-boys
categories: []
tags: 
    - monte-carlo
publishdate: '2020-02-29'
lastmod: '2020-02-29'
editor_options: 
  chunk_output_type: console
---

The origin of this question is an interview question from Google.

> In a village in which people only want boys every family continues to have children until they have a boy. If they have a girl, they have another child. If they have a boy, they stop. What is the expected boys to girls ratio in the village?

Though this can be solved mathematically (of course it can), the question is (IMO) easier to solve by thinking in a tree/branches structure.

If we assume the village has N families, and the probability of having a boy or girl is 50/50, then

```{r, echo=FALSE, fig.width=7, fig.align='center'}
DiagrammeR::mermaid("
    graph TB
        N --have boys, stop--> L21[N/2]
        N --have girls, continue--> L22[N/2]
        L22 --boys--> L221[N/4]
        L22 --girls--> L222[N/4]
        L222 --boys--> L2221[N/8]
        L222 --girls--> L2222[N/8]
        L2222 --> L222221[N/16]
        L2222 --> L222222[N/16]
        L222222 --- e1[...]
")
```

Here we see the obvious, as the depth of tree extends, the boy-to-girl disparity will ultimately become insignificant. The ratio will be very close to 50/50. 

For practice and validation sake, lets do a simulation.

```{r, collapse=TRUE}
library(magrittr)

simulate_once <- function(N, verbose=FALSE) {
    
    # nobody has anything yet...
    boys  = 0; girls = 0;
    
    print_vars <- function() {
        message("BGR:", boys, "/", girls)
        message("==================")
    }
    
    tryonemore <- function(x, msgON) {
        # browser()
        if(msgON) message(x)
        
        # terminal condition
        if (all(x)) {
            boys  <<- boys + sum(x)
            if (msgON) print_vars()
            return(1)
        }
        
        # assign to parent env so that we can keep track
        boys  <<- boys + sum(x)
        girls <<- girls + sum(!x)
        
        if (msgON) print_vars()
        return(tryonemore(rbinom(sum(!x), 1, 0.5), msgON = msgON) + sum(x))
    }
    # init call
    tryonemore(rbinom(N, 1, 0.5), msgON=verbose)
    return(boys / girls)
}
set.seed(1111)
simulate_once(10, verbose = TRUE)
```


```{r, cache=TRUE, message=FALSE}
library(purrr)
library(ggplot2)

res <- map_dbl(11:1000, ~ {
    mean(replicate(1000, simulate_once(.x)))
})

par(family = "Menlo")
plot(res,
     type  = "l",
     xlab  = "N (# families)",
     ylab  = "Expected Boys-to-Girls Ratio",)
title(main = "Simulate N from x1 to x2, 1000 times each", adj = 0)
dev.off()
```




