---
title: "EDA: A Survey into Market Risk Premium"
author: Thomas Jc
date: '2020-02-23'
slug: a-survey-into-market-risk-premium
categories: []
tags: 
    - finance
publishdate: '2020-02-23'
lastmod: '2020-02-23'
---

This is an exploratory analysis based on [the paper](https://ssrn.com/abstract=3358901) cited below. I am interested to explore the relationship of coefficient of variation (CV) between developed and developing countries.

> Fernandez, Pablo and Martinez, Mar and Fernández Acín, Isabel, Market Risk Premium and Risk-Free Rate Used for 69 Countries in 2019: A Survey (March 23, 2019). Available at SSRN: https://ssrn.com/abstract=3358901 or http://dx.doi.org/10.2139/ssrn.3358901

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(ggthemes)
library(ggforce)
library(janitor)
library(DT)

old <-
    theme_set(
        theme_economist_white(base_family = "Menlo") +
            theme(
                plot.background = element_rect(fill = "white"),
                legend.background = element_rect(fill = "white"), 
                legend.position = "bottom", 
                legend.key.height = unit(2, "mm"),
                legend.text = element_text(size = 7),
                title = element_text(size = 9)
            )
    )

# extract from pdf
raw <- read_csv("../etc/riskpremium/mrp2019.csv")
convert_to_numeric <- function(x) {
    x %>% 
        str_remove("%$") %>% 
        str_replace(",", ".") %>% 
        as.double()
}
mrp <- raw %>%
    clean_names() %>%
    mutate_at(vars(-contains("country")), convert_to_numeric)
```

To classify countries into developed or developing status, I use a general rough guide where country with GDP per capita above 24000 is considered developed. Alternatively, I could use human development index (HDI) but the former is commonly used among many economists. 

```{r, message=FALSE, echo=FALSE}
# from World Bank (year 2018)
raw <- read_csv("../etc/riskpremium/gdpcap2018.csv")
gdp <- raw %>% 
    clean_names() %>% 
    mutate(
        country = str_remove(country, "\\s$"),
        developed = if_else(gdp_cap2018 >= 24000, 1, 0)
    ) %>% 
    select(country, developed)
# this is our working data frame
df <- mrp %>% left_join(gdp, by = "country")
DT::datatable(df)
```

----

### EDA 

```{r}
hl.this = c("China", "United States of America", "Germany", "Japan")

df %>% 
    ggplot(aes(median, cv, col = log(answers))) +
    geom_point() + 
    scale_y_continuous(limits = c(0, 50)) +
    geom_mark_circle(aes(description = country, 
                         filter = country %in% hl.this),
                     col = "darkgray", con.colour = "darkgray") +
    labs(x     = "Median of MRP", 
         y     = "Coefficient of Variation", 
         col   = "Sample Size (log)",
         title = "Market Risk Premium: A Survey (Fernandez, 2019)")
```

Developed or not does not seem to exhibit any relationship to variation.

```{r}
df %>% 
    filter(!is.na(developed)) %>% 
    ggplot(aes(factor(developed), cv)) + 
    geom_point(position = position_jitter(width = 0.2, height = 0)) + 
    scale_x_discrete(labels = c(`0` = "false", `1` = "true")) +
    labs(x = "Developed Country", y = "Coefficient of Variation")
```



