---
title: STM Package Walkthrough Part Two
author: tmasjc
date: '2020-04-05'
slug: stm-package-walkthrough-part-two
categories: []
tags: 
    - topic-model
    - stm
publishdate: '2020-04-05'
lastmod: '2020-04-05'
---


```{r, echo=FALSE}
dat <- read.csv("~/Downloads/poliblogs2008.csv", stringsAsFactors = FALSE)[-1]
```

This is our working data.

```{r}
library(stm)
dplyr::glimpse(dat)
```

First ingest and prepare documents.

```{r, cache=TRUE}
processed <- textProcessor(dat$documents, metadata = dat)
out <-
    prepDocuments(processed$documents,
                  processed$vocab,
                  processed$meta,
                  lower.thresh = 10)
```

`stm` package has a built-in function `searchK` to assist in selecting the best `K` parameter for topic modeling. 

```{r, eval=FALSE}
storage <- searchK(
    documents  = out$documents,
    vocab      = out$vocab,
    K          = c(15:25),
    cores      = 8, ## parallel computing
    prevalence =  ~ rating + s(day), 
    data       = out$meta
)
```

```{r, echo=FALSE}
load("../etc/stm/storage.RData")
```

Measures:

1. *Semantic Coherence* - related to pointwise mutual information (Mimno et al 2011). In models that are semantically coherent, words which are most probable under a topic should co-occur within the same document. 

2. *Exclusivity* - based on FREX, weighted harmonic mean of the word's rank in terms of exclusivity and frequency.

3. *Heldout Likelihood* - hold out some fraction of the words in training and use the document-level latent variables to evaluate the probability of the heldout portion.

4. *Bound* - The change in the approximate variational lower bound to convergence.

5. *Residuals* - multinomial dispersion of the STM residuals $\sigma^2 = 1$. If we calculate the sample dispersion and the value is greater than one, this implies that $K$ is set too low.

```{r}
storage$results
```


```{r, message=FALSE}
library(dplyr)
library(ggplot2)
library(ggforce)

old <-
    theme_set(theme_minimal(base_family = "Menlo") + 
                  theme(axis.text.x = element_text(angle = 90)))
df  <- storage$results %>% mutate(K = as.factor(K))

df %>% 
    ggplot(aes(K, semcoh, group = 1)) +
    geom_line() +
    labs(x = "", y = "", title = "Semantic Coherence")

df %>% 
    ggplot(aes(K, heldout, group = 1)) + 
    geom_line() + 
    labs(x = "", y = "", title = "Heldout Prob.")

df %>% 
    ggplot(aes(K, residual, group = 1)) + 
    geom_line() +
    labs(x = "", y = "", title = "Residuals")

df %>% 
    ggplot(aes(semcoh, exclus, col = K)) +
    geom_point(size = 3) +
    geom_mark_circle(aes(filter = K %in% c(17, 19, 22)), 
                     col = "red", description = "Potentials best candidates for K") +
    labs(x = "Semantic Coherence", y = "Exclusivity") +
    theme(legend.position = "bottom")
```

We have narrowed the range of $K$ and may proceed to do further inspection via other techniques.